### 2019-07-02 2-Chapters/3-Chapter/chapter3.tex
Lots of things to do, oh my god.
Done for Chapters 2 and 3

\chapter{SMPyBandits: a state-of-the-art Python library to simulate MAB problems}
\label{chapter:3}

@@

Since December $2016$ I worked regularly on the SMPyBandits library, which is a Python package  designed to allow easy and fast numerical simulations on single-player and multi-players Multi-Armed Bandits (MAB) algorithms.
%
This library is by far the most exhaustive open-source implementation of state-of-the-art algorithms and different kinds of MAB models.
It aims at being exhaustive, simple to use and maintain, and has a clean and well documented codebase, and it uses the popular open-source Python language.
It allows fast prototyping of simulations, with an easy configuration system and command-line options to customize experiments.
Experiment results are saved in an optimized binary format (HDF5) as well as high quality plots of many useful visualizations.
%
More than two years of active development have shown how easy it can be to add new algorithms, new arm distributions, and new bandit models (\eg, Markov or non-stationary).
It is hosted on GitHub, and uses the state-of-the-art development technologies, by using two online services of continuous integration to run automated tests and to build its online documentation.

SMPyBandits stands for \emph{S}ingle- and \emph{M}ulti-\emph{P}layer \emph{Bandits} in P\emph{y}thon.
The library does not aim at being blazing fast or optimized in terms of memory usage, as it comes with a pure Python implementation \cite{python}, and uses only standard open-source Python packages (\ie, no hard to install dependencies).
Some critical parts are also available as a \texttt{C} Python extension, and the just-in-time Numba compiler \cite{numba} is used whenever it is possible, so we can note that we optimized the time efficiency of what could be (easily) optimized.
However if simulation speed really matters, one should rather refer to less exhaustive but faster implementations, like for example \cite{TorLibbandit} in \texttt{C++} or \cite{VishMABjl} in Julia. Note that both are not maintained anymore, and contain just a few algorithms for the simple stationary MAB model.

In this Chapter~\ref{chapter:3}, we start by presenting in Section~\ref{sec:3:presentationLibrary} the organization of the library.
We use the library to compare the most famous and most efficient single-player MAB algorithms in Section~\ref{sec:3:reviewSPAlgorithms},
then we discuss in Section~\ref{sec:3:timeAndMemoryCosts} about the time and memory costs of MAB algorithms.
The take away messages are two-fold:
from a practical point-of-view, it is usually advised to use the simplest algorithm and we advise to use \UCB{} (as we do in the next Chapter~\ref{chapter:4}),
and that for theoretical works where optimality of the MAB algorithm matters, we advise to use \klUCB{} (as we do in the next Chapters~\ref{chapter:5} and \ref{chapter:6}).
% To also illustrate how one can easily extend the SMPyBandits library to add not only add new distributions or new algorithms.
% but also new bandit models, we detail in Section~\ref{sec:3:markovModels} how we added the possibility to simulate Markovian models (as introduced by \cite{Anantharam87b}).
%
Finally, in Appendix~\ref{sec:3:appendix} we include some small files showing how to use the library, and we conclude with some details regarding the use of parallel computing to speed-up simulations.

@@

We start by explaining how SMPyBandits simulates MAB problems, by detailing its components:
MAB problems (single- or multi-players),
and reward distributions.
%
We then detail how it implements MAB algorithms (\eg, \UCB),
and what kind of information is displayed, saved and plotted after a batch of simulations of bandit problems.
%
%
The main goal of the library is to be easily able to simulate MAB algorithms (\eg, three algorithms like \UCB, \klUCB{} and Thompson sampling) on one or more bandit models (single- or multi-players) defined by the number of arms $K$ and their distributions $(\nu_k)_{k\in[K]}$, for some time steps that range from $t=1$ to the horizon $T$.
After the simulation, the library then displays statistical summary of the (mean) rewards accumulated by each algorithm, as well as regret and other visualizations.
The same problem is simulated for $N$ independent repetitions (\eg, $N=100$) in order to show mean results with a low variance.

@@

For the classical single-player stochastic MAB model, as defined in Chapter~\ref{chapter:2},
a stochastic MAB problem is defined by $K>1$ distributions $(\nu_k)_{k\in[K]}$ (also called arms),
used to generate the \iid{} samples $Y_{k,t} \sim \nu_k, \forall t$.
%
An agent chooses arm $A(t)\in[K]$ at time $t$ and observes the reward $r(t) = Y_{A(t),t}$ without knowing the other (hidden) rewards (in opposition to the full-information setting where $(Y_{1,t},\dots,Y_{K,t})$ is observed)).
%
Her goal is to maximize $\sum_{t=1}^T r(t)$, which require to trade off between exploring the unknown $K$ arms, exploiting the arm which is currently estimated to be the best one.

\textbf{Simulation loop for single-player MAB.}
%
Any simulation library targeting single-player bandit problems must implement at least three components:
reward distributions, MAB algorithm, and a simulation loop that essentially looks like this.
% \begin{itemize}
    % \item
    First, initialize the MAB problem and one or more algorithms,
    % \item
    Then, for $t=1$ to $t=T$ (known before hand), repeat the following block (for each algorithm). Ask algorithm $\cA$ her chosen arm $A(t)$, then sample a (random) reward $r(t)$ (\iid) from distribution $\nu_{A(t)}$, and finally feeds the observation $(A(t), r(t))$ to the algorithm,
    % \item
    At the end, compute the cumulated reward, the regret, plot visualizations etc.
% \end{itemize}
%
Note that the second step is repeated a large number of times (\eg, $N=100$), in order to study \emph{mean} regret and not only the regret in one single trajectory.
If one wants to compare algorithms on a same problem, it is possible to sample all the rewards $(Y_{k,t})_{k\in[K], t\in[T]}$ before-hand, and store them so that for each repetition of the simulation, the randomness from the environment (\ie, the rewards) has the same impact on all the algorithms.

\textbf{Simulation loop for multi-players MAB.}
%
For Cognitive Radio dealing with OSA problems and other applications, a well-studied extension is to consider $M\geq2$ players, interacting on the same $K$ arms.
Whenever two or more players select the same arm at the same time (\eg, for OSA, the same frequency channel), they all suffer from a radio collision.
%
Different collision models have been proposed, and the simplest one consists in giving a $0$ reward to each colliding players.
Without any centralized supervision or coordination between players, they must learn to access the $M$ best resources (\ie, arms with highest means) without collisions.
We refer to Chapter~\ref{chapter:5} which introduces the multi-players bandit model.

SMPyBandits implements all collision models found in the literature (in the module \texttt{\href{https://SMPyBandits.GitHub.io/docs/Environment.CollisionModels.html}{Environment.CollisionModels}}), as well as all the algorithms known by the authors (in the module \texttt{\href{https://SMPyBandits.GitHub.io/docs/PoliciesMultiPlayers.html}{PoliciesMultiPlayers}}).
%
In particular, it includes
\texttt{rhoRand}
% (\texttt{\href{https://SMPyBandits.GitHub.io/docs/PoliciesMultiPlayers.rhoRand.html}{PoliciesMultiPlayers.rhoRand}})
from \cite{Anandkumar11},
\texttt{MEGA}
% (\texttt{\href{https://SMPyBandits.GitHub.io/docs/Policies.MEGA.html}{Policies.MEGA}})
from \cite{Avner15},
\texttt{MusicalChair}
% (\texttt{\href{https://SMPyBandits.GitHub.io/docs/Policies.MusicalChair.html}{Policies.MusicalChair}})
from \cite{Rosenski16},
and our state-of-the-art algorithms
\texttt{RandTopM}
% (\texttt{\href{https://SMPyBandits.GitHub.io/docs/PoliciesMultiPlayers.RandTopM.html}{PoliciesMultiPlayers.RandTopM}})
and \texttt{MCTopM}
% (\texttt{\href{https://SMPyBandits.GitHub.io/docs/PoliciesMultiPlayers.MCTopM.html}{PoliciesMultiPlayers.MCTopM}})
from \cite{Besson2018ALT}, and more recent solutions.
For comparison, realistic (\eg, \UCB{} for multiple play) or full-knowledge \emph{centralized} algorithms are also implemented.

Any simulation library targeting multi-players bandit problems must implement at least another component:
% a collision model, and
a simulation loop that essentially looks like this.
% \begin{itemize}
    % \item
    First, initialize the MAB problem with $M$ players, and one or more cohorts of $M$ players (one player is one algorithm, usually $M$ times the same one),
    % \item
    For $t=1$ to $t=T$ (known before hand), repeat the following block (for each algorithm). Ask each player $\cA^j$ her chosen arm $A^j(t)$, then query the collision model to know which player will get a zero reward (for a collision) and which player will get a random reward from the environment.
    Then, sample (random) feedback $Y_{k,t}$ (\iid) from distributions $\nu_{k}$, and compute the rewards $r^j(t)$ from the random feedback using the collision model.
    Finally, feed the observation $(A^j(t), Y_{A^j(t),t}, r^j(t))$ to player $j$, for all the $M$ players,
    %
    The default collision model is the most widely studied in the literature, where a player encounters a collision (\ie, received a zero reward $r^j(t)=0$) if she is not the only one to have chosen an arm $k=A^j(t)$, otherwise $r^j(t)=Y_{A^j(t),t}$ if she is the only one playing this arm.
    % \item
    At the end, compute the accumulated reward, the regret, plot visualizations etc.
% \end{itemize}

\subsection{Reward distributions}
%
We focus on one dimensional distributions, implemented in the \texttt{\href{https://smpybandits.github.io/docs/Arms.html}{Arms}} module.
The library supports discrete distributions: \emph{Bernoulli} (\texttt{\href{https://smpybandits.github.io/docs/Arms.Bernoulli.html}{Bernoulli}}), \emph{binomial} (\texttt{\href{https://smpybandits.github.io/docs/Arms.Binomial.html}{Binomial}}), \emph{Poisson} (\texttt{\href{https://smpybandits.github.io/docs/Arms.Poisson.html}{Poisson}}), and a generic \emph{discrete} distribution (\texttt{\href{https://smpybandits.github.io/docs/Arms.DiscreteArm.html}{DiscreteArm}}),
as well as continuous distributions,
which can be truncated to an interval $[a,b]$ or have unbounded support ($\mathbb{R}$):
\emph{exponential} (\texttt{\href{https://smpybandits.github.io/docs/Arms.Exponential.html}{Exponential}}), \emph{gamma} (\texttt{\href{https://smpybandits.github.io/docs/Arms.Gamma.html}{Gamma}}), \emph{Gaussian} (\texttt{\href{https://smpybandits.github.io/docs/Arms.Gaussian.html}{Gaussian}}) and \emph{uniform} (\texttt{\href{https://smpybandits.github.io/docs/Arms.Uniform.html}{Uniform}}).
%
The default is to use the same distribution for the $K$ arms, as this is the setting studied in the literature,
but it also possible to mix them.
%  (\eg, one arm is Bernoulli-distributed and another).
%  (see for instance Figure~\ref{fig:25:HarderMixed}).
%
For instance the following code in Code~\ref{lst:3:pythonCodeCreateProblem} creates three arms, following a Bernoulli distribution with respective means $0.1, 0.5, 0.9$, and a \texttt{MAB} object which encapsulates the problem.
We give another example for \emph{truncated} Gaussian distributions (on $[0,1]$), and a visualization of a histogram of $10000$ rewards obtained from the arms, below in Figure~\ref{fig:3:exampleOfRewards}.

@@

We have not yet added support for higher dimensional distributions of rewards, \eg, linear bandit, but it would be an interesting extension of SMPyBandits.
%
However, note that our library does support finite-state real-valued Markov MAB models, where arms correspond to Markov chains \cite{Norris98}, in the two \emph{rested}/\emph{restless} variants, as introduced by \cite{Anantharam87b}.

@@

SMPyBandits is a complete open-source implementation of single-player (classical) bandit algorithms,
containing over 65 algorithms (in the module \texttt{\href{https://SMPyBandits.GitHub.io/docs/Policies.html}{Policies}}).
It uses a well-designed hierarchical structure and class inheritance scheme (as detailed on the various UML diagrams shown on the \texttt{\href{https://SMPyBandits.GitHub.io/uml_diagrams/README.html}{uml\_diagrams}} folder) to minimize redundancy in the codebase.
For instance, most existing algorithms are index policies (see Algorithm~\ref{algo:2:indexPolicy}), and new ones are easily written by inheriting from the \texttt{IndexPolicy} class (\texttt{\href{https://SMPyBandits.GitHub.io/docs/Policies.IndexPolicy.html}{Policies.IndexPolicy}}).
They compute an index $I_k(t)\in\mathbb{R}$ for each arm $k$ at time $t$, and play $A(t) = \arg\max_k I_k(t)$.
For instance the code specific to $\UCB_1$ is as short as this (and fully documented):

% https://tex.stackexchange.com/a/12430/
\begin{small}
% \begin{listing}[h!]
    \inputminted[linenos=true,numbersep=5pt,frame=lines,framesep=2mm]{python3}{2-Chapters/3-Chapter/src/example_of_a_IndexPolicy_UCB.py}
    % \caption{Small snippet of code defining the UCB algorithm, as a simple example of an Index Policy.}
    \captionof{listing}{Code defining the UCB algorithm, as a simple example of an Index Policy\label{lst:3:smallIndexPolicy}.}
    % \label{lst:3:smallIndexPolicy}
% \end{listing}
\end{small}

@@

\subsection{Documentation}

A complete documentation, for each algorithm and the entire codebase, is available online at
% even including the constants in the different configuration files:
\texttt{\href{https://SMPyBandits.GitHub.io}{SMPyBandits.GitHub.io}}.
It uses the Sphinx software \cite{sphinx}, and the content is directly written in the Python files as docstrings, so users have access to the documentation from within their IDE or the Python console.
The most interesting component of the library is the many MAB algorithms being implemented: for each of them, the documentation gives a reference to a research paper (\eg, \cite{Kaufmann12BUCB} for \texttt{BayesUCB}), as well as a bird-eye view of its behavior.
For most algorithms, especially for index policies, the internal variables of the implementation are carefully linked to the notations of each paper, and formulas explaining the way the algorithm selects its arm are usually given.
Whenever it is needed, we also included warnings or information about the empirical performance of the more costly algorithms.

@@

\subsection{Dependencies}
\label{sub:3:dependencies}

@@

Optional dependencies include \texttt{joblib} \cite{joblib} for parallel simulations, \texttt{numba} \cite{numba} for automatic speed-up on small functions using a Just-in-Time compiler, \texttt{sphinx} \cite{sphinx} for generating the documentation, \texttt{virtualenv} \cite{virtualenv} for launching simulations in isolated environments, and \texttt{jupyter} \cite{jupyter} used with \texttt{ipython} \cite{ipython} to experiment with the code.

@@

Since $2017$ and $2018$, SMPyBandits is using two online continuous integration (CI) service, to automatically build and host the documentation, and to automatically test the library on some numerical simulations.
%
These CI services are free for open-source works, and are both triggered whenever a modification of the codebase is sent to the hosting platform GitHub.

@@

\textbf{Travis CI.}
Since $2018$, we use the free web service provided by Travis CI (\href{https://travis-ci.org/}{\texttt{Travis-CI.org}}).
Travis CI allows to automatically run short numerical simulations after every commit.
This allows to check that each modification on any part of the codebase does not break anything, as well as giving an up-to-date example of log files that shows online the results of different examples of experiments. The tests covers all the main models and almost all the algorithms implemented in SMPyBandits, and it has been proven very useful to quickly find and fix new bugs.
For more details, see \href{https://travis-ci.org/SMPyBandits/SMPyBandits}{\texttt{Travis-CI.org/SMPyBandits/SMPyBandits}}.
%
Since its first use, the service did about 160 builds, and about $30\%$ of them were useful to detect newly introduced issues or bugs in the code.
The build is configured using the \texttt{.travis.yml} file in SMPyBandits main folder, and it runs numerical experiments with a short horizon (\eg, $T=100$) and a small number of repetitions (\ie, $N=4$), but covering all the different models implemented by SMPyBandits (and even including some that are not covered in this chapter, like the Markov model, or the sparse multi-players model).
Builds typically run for $15$ minutes, and Travis CI have been proven to be very useful in our development process.

@@

We focused here on the most well known algorithms, discussed in Section~\ref{sec:2:famousMABalgorithms}, and two others that are efficient but less known (AdBandits and BESA).
To highlight that SMPyBandits implements many other MAB algorithms proposed in recent works, we detail in Appendix~\ref{sub:3:additionalExperiments} $7$ other algorithms.
They are more recent (\eg, $3$ from $2018$ and $2$ from early $2019$) and are usually more complex, or are based on a different point-of-view.
The numerical results presented below include these extra algorithms, to justify empirically that despite their respective qualities, and despite being very efficient, it is reasonable to focus on \UCB{} and \klUCB{} in the rest of this thesis, as they stay comparable with the state-of-the-art algorithms but are simpler to implement (\eg, when comparing \UCB{} to \texttt{ApproximatedFHGittins} or $\mathrm{MOSS}$-$\mathrm{Anytime}$) to manipulate theoretically (\eg, when comparing \klUCB{} to BESA, PHE or RCB).

@@

$(ii)$
We detailed how SMPyBandits is implementing arms, problems, algorithms, and use these components to implement a simulation loop, with various visualizations being performed after the simulation.
As far as now, SMPyBandits is restricted to the finite-arm cases, but it supports a wide range of arms distributions.
Different kinds of models are implemented, from stationary single-player to piece-wise stationary multi-players with different collision models.
One of the main qualities of the library is that it is quite exhaustive, as all the main families of algorithms covering these different models have been implemented, even very recent algorithms from the literature, as we tried our best in following the active research since December $2016$ to June $2019$.
More than $65$ algorithms or variants of algorithms are implemented for the single-player case, $5$ for the expert aggregation problem, about $15$ for the multi-players case, and about $20$ for the piece-wise stationary problem.
All the codebase is fully documented, and the library is using continuous integration to run automated tests on the code after every modification.
%
When comparing algorithms on a problem, the main performance measure is the regret, but the library also computes, stores and visualizes other measures, such as best-arm selection rate or mean cumulated reward, as well as real time and memory costs.

>>>>

% \chapter{SMPyBandits: a state-of-the-art Python library to simulate MAB problems}
\chapter{SMPyBandits: our Python library to simulate MAB problems}
\label{chapter:3}

@@

Since December $2016$ I worked regularly on the SMPyBandits library, which is a Python package  designed to allow easy and fast numerical simulations on single-player and multi-players Multi-Armed Bandits (MAB) algorithms.
%
This library is by far the most exhaustive open-source implementation of state-of-the-art algorithms and different kinds of MAB models.
It aims at being exhaustive, simple to use and maintain, and has a clean and well documented codebase, and it uses the popular open-source Python language.
It allows fast prototyping of simulations, with an easy configuration system and command-line options to customize experiments.
Experiment results are saved in an optimized binary format (HDF5) as well as high quality plots of useful visualizations.
%
More than two years of active development have shown how easy it can be to add new algorithms, new arm distributions, and new bandit models (\eg, Markov or non-stationary).
It is hosted on GitHub, and uses the state-of-the-art development technologies, by using two online services of continuous integration to run automated tests and to build its online documentation.

SMPyBandits stands for \emph{S}ingle- and \emph{M}ulti-\emph{P}layer \emph{Bandits} in P\emph{y}thon.
The library does not aim at being blazing fast or optimized in terms of memory usage, as it comes with a Python implementation \cite{python}, and uses only standard open-source Python packages (\ie, no hard to install dependencies).
Some critical parts are also available as a \texttt{C} Python extension, and the just-in-time Numba compiler \cite{numba} is used whenever it is possible, so we can note that we optimized the time efficiency of what could be (easily) optimized.
However if simulation speed really matters, one should rather refer to less exhaustive but faster implementations, like for example \cite{TorLibbandit} in \texttt{C++} or \cite{VishMABjl} in Julia. Note that both are not maintained anymore, and contain just a few algorithms for the simple stationary MAB model.

In this Chapter~\ref{chapter:3}, we start by presenting in Section~\ref{sec:3:presentationLibrary} the organization of the library.
We use the library to compare the most famous and most efficient single-player MAB algorithms in Section~\ref{sec:3:reviewSPAlgorithms},
then we discuss in Section~\ref{sec:3:timeAndMemoryCosts} about the time and memory costs the same algorithms.
The take away messages are two-fold:
from a practical point-of-view, it is usually advised to use the simplest algorithm and we advise to use \UCB{} (as we do in the next Chapter~\ref{chapter:4}),
and that for theoretical works where optimality of the MAB algorithm is analyzed, we advise to use \klUCB{} (as we do in the next Chapters~\ref{chapter:5} and \ref{chapter:6}).
% To also illustrate how one can easily extend the SMPyBandits library to add not only add new distributions or new algorithms.
% but also new bandit models, we detail in Section~\ref{sec:3:markovModels} how we added the possibility to simulate Markovian models (as introduced by \cite{Anantharam87b}).
%
Finally, in Appendix~\ref{sec:3:appendix} we include some small files showing how to use the library, and we conclude with some details regarding the use of parallel computing to speed-up simulations.

@@

We start by explaining how SMPyBandits simulates MAB problems, by detailing its components:
MAB problems (single- or multi-players),
and reward distributions.
%
We then detail how it implements MAB algorithms (\eg, \UCB),
and what kind of information is displayed, saved and plotted after a batch of simulations of bandit problems.
%
%
The main goal of the library is to be easily able to simulate MAB algorithms (\eg, three algorithms like \UCB, \klUCB{} and Thompson sampling) on one or more bandit models (single- or multi-players) defined by the number of arms $K$ and their distributions $(\nu_k)_{k\in[K]}$, for some time steps that range from $t=1$ to the horizon $T$.
After the simulation, the library then displays statistical summary of the (mean) rewards accumulated by each algorithm, as well as regret and other visualizations.
The same problem is simulated for $N$ independent repetitions (\eg, $N=100$) in order to show mean results with low variances.

@@

For the classical single-player stochastic MAB model, as defined in Chapter~\ref{chapter:2},
a stochastic MAB problem is defined by $K \geq 2$ distributions $(\nu_k)_{k\in[K]}$ (also called arms),
used to generate the \iid{} samples $Y_{k,t} \sim \nu_k, \forall t$.
%
An agent chooses arm $A(t)\in[K]$ at time $t$ and observes the reward $r(t) = Y_{A(t),t}$ without knowing the other (hidden) rewards (in opposition to the full-information setting where $(Y_{1,t},\dots,Y_{K,t})$ is observed).
%
Her goal is to maximize $\sum_{t=1}^T r(t)$, which require to trade off between exploring the unknown $K$ arms, exploiting the arm which is currently estimated to be the best one.

\paragraph{Simulation loop for single-player MAB.}
%
Any simulation library targeting single-player bandit problems must implement at least three components:
reward distributions, MAB algorithms, and a simulation loop that essentially looks like this.
% \begin{itemize}
    % \item
    First, initialize the MAB problem and one or more algorithms,
    % \item
    Then, for $t=1$ to $t=T$ (known before hand), repeat the following block (for each algorithm). Ask algorithm $\cA$ her chosen arm $A(t)$, then sample a (random) reward $r(t)$ (\iid) from distribution $\nu_{A(t)}$, and finally feeds the observation $(A(t), r(t))$ to the algorithm.
    % \item
    At the end, compute the cumulated reward, the regret, then plot visualizations etc.
% \end{itemize}
%
Note that the second step should be repeated a large number of times (\eg, $N=100$), in order to study \emph{mean} regret and not only the regret in one single trajectory.
If one wants to compare algorithms on a same problem, it is possible to sample all the rewards $(Y_{k,t})_{k\in[K], t\in[T]}$ before-hand, and store them so that for each repetition of the simulation, the randomness from the environment (\ie, the rewards) has the same impact on all the algorithms (with the option \texttt{configuration["cache\_reward"] = True}).

\paragraph{Simulation loop for multi-players MAB.}
%
For Cognitive Radio dealing with OSA problems and other applications, an extension is to consider $M\geq2$ players, interacting on the same $K$ arms.
Whenever two or more players select the same arm at the same time (\eg, for OSA, the same frequency channel), they all suffer from a radio collision.
%
Different collision models have been proposed, and the simplest one consists in giving a $0$ reward to each colliding players.
Without any centralized supervision or coordination between players, they must learn to access the $M$ best resources (\ie, arms with highest means) without collisions.
We refer to Chapter~\ref{chapter:5} which introduces the multi-players bandit model.

SMPyBandits implements all the collision models found in the literature (in the module \texttt{\href{https://SMPyBandits.GitHub.io/docs/Environment.CollisionModels.html}{Environment.CollisionModels}}), as well as all the algorithms known by the authors (in the module \texttt{\href{https://SMPyBandits.GitHub.io/docs/PoliciesMultiPlayers.html}{PoliciesMultiPlayers}}).
%
In particular, it includes
\texttt{rhoRand}
% (\texttt{\href{https://SMPyBandits.GitHub.io/docs/PoliciesMultiPlayers.rhoRand.html}{PoliciesMultiPlayers.rhoRand}})
from \cite{Anandkumar11},
\texttt{MEGA}
% (\texttt{\href{https://SMPyBandits.GitHub.io/docs/Policies.MEGA.html}{Policies.MEGA}})
from \cite{Avner15},
\texttt{MusicalChair}
% (\texttt{\href{https://SMPyBandits.GitHub.io/docs/Policies.MusicalChair.html}{Policies.MusicalChair}})
from \cite{Rosenski16},
and our state-of-the-art algorithms
\texttt{RandTopM}
% (\texttt{\href{https://SMPyBandits.GitHub.io/docs/PoliciesMultiPlayers.RandTopM.html}{PoliciesMultiPlayers.RandTopM}})
and \texttt{MCTopM}
% (\texttt{\href{https://SMPyBandits.GitHub.io/docs/PoliciesMultiPlayers.MCTopM.html}{PoliciesMultiPlayers.MCTopM}})
from \cite{Besson2018ALT}, and more recent solutions.
For comparison, realistic (\eg, \UCB{} for multiple play) or full-knowledge \emph{centralized} algorithms are also implemented.

Any simulation library targeting multi-players bandit problems must implement at least another component:
% a collision model, and
a simulation loop that essentially looks like this.
% \begin{itemize}
    % \item
    First, initialize the MAB problem with $M$ players, and one or more cohorts of $M$ players (one player is one algorithm, usually $M$ times the same one),
    % \item
    For $t=1$ to $t=T$ (known before hand), repeat the following block (for each cohort). Ask each player $\cA^j$ her chosen arm $A^j(t)$, then query the collision model to know which player will get a zero reward (for a collision) and which player will get a random reward from the environment.
    Then, sample (random) feedback $Y_{k,t}$ (\iid) from distributions $\nu_{k}$, and compute the rewards $r^j(t)$ from the random feedback using the collision model.
    Finally, feed the observation $(A^j(t), Y_{A^j(t),t}, r^j(t))$ to player $j$, for all the $M$ players,
    %
    The default collision model is the most widely studied in the literature, where a player encounters a collision (\ie, received a zero reward $r^j(t)=0$) if she is not the only one to have chosen an arm $k=A^j(t)$, otherwise $r^j(t)=Y_{A^j(t),t}$ if she is the only one playing this arm.
    % \item
    At the end, compute the accumulated reward, the regret, then plot visualizations etc.
% \end{itemize}

\subsection{Reward distributions}
%
We focus on one dimensional distributions (\ie, $r(t)\in\R$), implemented in the \texttt{\href{https://smpybandits.github.io/docs/Arms.html}{Arms}} module.
The library supports discrete distributions: \emph{Bernoulli} (\texttt{\href{https://smpybandits.github.io/docs/Arms.Bernoulli.html}{Bernoulli}}), \emph{binomial} (\texttt{\href{https://smpybandits.github.io/docs/Arms.Binomial.html}{Binomial}}), \emph{Poisson} (\texttt{\href{https://smpybandits.github.io/docs/Arms.Poisson.html}{Poisson}}), and a generic \emph{discrete} distribution (\texttt{\href{https://smpybandits.github.io/docs/Arms.DiscreteArm.html}{DiscreteArm}}),
as well as continuous distributions,
which can be truncated to an interval $[a,b]$ or have unbounded support ($\mathbb{R}$):
\emph{exponential} (\texttt{\href{https://smpybandits.github.io/docs/Arms.Exponential.html}{Exponential}}), \emph{gamma} (\texttt{\href{https://smpybandits.github.io/docs/Arms.Gamma.html}{Gamma}}), \emph{Gaussian} (\texttt{\href{https://smpybandits.github.io/docs/Arms.Gaussian.html}{Gaussian}}) and \emph{uniform} (\texttt{\href{https://smpybandits.github.io/docs/Arms.Uniform.html}{Uniform}}).
%
The default is to use the same distribution for the $K$ arms, as this is the setting studied in the literature,
but it also possible to mix them.
%  (\eg, one arm is Bernoulli-distributed and another).
%  (see for instance Figure~\ref{fig:25:HarderMixed}).
%
For instance the following code in Code~\ref{lst:3:pythonCodeCreateProblem} creates three arms, following Bernoulli distributions, with respective means $0.1, 0.5, 0.9$, and a \texttt{MAB} object which encapsulates the problem.
We give another example for \emph{truncated} Gaussian distributions on $[0,1]$, and a visualization of a histogram of $10000$ rewards, below in Figure~\ref{fig:3:exampleOfRewards}.

@@

We have not yet added support for higher dimensional distributions of rewards, such as linear bandits, but it would be an interesting extension of SMPyBandits.
%
However, note that our library does support finite-state real-valued Markov MAB models, where arms correspond to Markov chains \cite{Norris98}, in the two \emph{rested}/\emph{restless} variants, as introduced by \cite{Anantharam87b}.

@@

SMPyBandits is a complete open-source implementation of single-player (classical) bandit algorithms,
containing over 65 algorithms (in the module \texttt{\href{https://SMPyBandits.GitHub.io/docs/Policies.html}{Policies}}).
It uses a well-designed hierarchical structure and class inheritance scheme (as detailed on the various UML diagrams shown on the \texttt{\href{https://SMPyBandits.GitHub.io/uml_diagrams/README.html}{uml\_diagrams}} folder) to minimize redundancy in the codebase.
For instance, most existing algorithms are index policies (see Algorithm~\ref{algo:2:indexPolicy}), and new ones are easily written by inheriting from the \texttt{IndexPolicy} class (\texttt{\href{https://SMPyBandits.GitHub.io/docs/Policies.IndexPolicy.html}{Policies.IndexPolicy}}).
% They compute an index $U_k(t)\in\mathbb{R}$ for each arm $k$ at time $t$, and play $A(t) \in \arg\max_k U_k(t)$.
For instance the code specific to $\UCB_1$ (with $\alpha=2$) is as written and fully documented in a file as short as this:

% https://tex.stackexchange.com/a/12430/
\begin{small}
% \begin{listing}[h!]
    \inputminted[linenos=true,numbersep=5pt,frame=lines,framesep=2mm]{python3}{2-Chapters/3-Chapter/src/example_of_a_IndexPolicy_UCB.py}
    % \caption{Small snippet of code defining the UCB algorithm, as a simple example of an Index Policy.}
    \captionof{listing}{Code defining the $\UCB_1$ algorithm, as a simple example of an Index Policy\label{lst:3:smallIndexPolicy}.}
    % \label{lst:3:smallIndexPolicy}
% \end{listing}
\end{small}

@@

\subsection{Documentation of the library}

A complete documentation, for each algorithm and the entire codebase, is available online at
% even including the constants in the different configuration files:
\texttt{\href{https://SMPyBandits.GitHub.io}{SMPyBandits.GitHub.io}}.
It uses the Sphinx software \cite{sphinx}, and the content is directly written in the Python files as docstrings (in \texttt{"""\dots"""}, see in the example of Code~\ref{lst:3:smallIndexPolicy} given above), so users have access to the documentation from within their IDE or the Python console.
The most interesting component of the library is the many MAB algorithms being implemented: for each of them, the documentation gives a reference to a research paper (\eg, \cite{Auer02} for \texttt{UCB} in Code~\ref{lst:3:smallIndexPolicy}), as well as a bird-eye view of its behavior.
For most algorithms, especially for index policies, the internal variables of the implementation are carefully linked to the notations of each paper, and formulas explaining the way the algorithm selects its arm are usually given.
Whenever it is needed, we also included warnings or information about the empirical performance of the more costly algorithms.

@@

\subsection{Dependencies on other Python libraries}
\label{sub:3:dependencies}

@@

Optional dependencies include \texttt{joblib} \cite{joblib} for parallel simulations, \texttt{numba} \cite{numba} for automatic speed-up on small functions (using a Just-in-Time compiler), \texttt{sphinx} \cite{sphinx} for generating the documentation, \texttt{virtualenv} \cite{virtualenv} for launching simulations in isolated environments, and \texttt{jupyter} \cite{jupyter} used with \texttt{ipython} \cite{ipython} to experiment with the code.

@@

Since $2017$ and $2018$, SMPyBandits is using two online continuous integration (CI) services, to automatically build and host its documentation, and to automatically test the library on some numerical simulations.
%
These CI services are free for open-source works, and are both triggered whenever a modification of the codebase is sent to the hosting platform GitHub.

@@

\textbf{Travis CI.}
Since $2018$, we use the free web service provided by Travis CI (\href{https://travis-ci.org/}{\texttt{Travis-CI.org}}).
It allows to automatically run short numerical simulations after every commit,
in order to check that each modification on any part of the codebase does not break anything, as well as giving an up-to-date example of log files that shows online the results of different examples of experiments. The tests covers all the main models and almost all the algorithms implemented in SMPyBandits, and it has been proven very useful to quickly find and fix new bugs.
For more details, see \href{https://travis-ci.org/SMPyBandits/SMPyBandits}{\texttt{Travis-CI.org/SMPyBandits/SMPyBandits}}.
%
Since its first use, the service did about $190$ builds, and about $30\%$ of them were useful to detect newly introduced issues or bugs in the code.
The build is configured using the \texttt{.travis.yml} file in SMPyBandits main folder, and it runs numerical experiments with a short horizon (\eg, $T=100$) and a small number of repetitions (\ie, $N=4$), but covering all the different models implemented by SMPyBandits (and even including some that are not covered in this chapter, like the Markov model, or the sparse multi-players model).
Builds typically run for $15$ minutes, and Travis CI have been proven to be very useful in our development process.

@@

We focused here on the most well known algorithms, discussed in Section~\ref{sec:2:famousMABalgorithms}, and two others that are efficient but less known (AdBandits and BESA).
To highlight that SMPyBandits implements many other MAB algorithms proposed in recent works, we detail in Appendix~\ref{sub:3:additionalExperiments} $7$ other algorithms.
They are more recent (\eg, $3$ from $2018$ and $2$ from early $2019$) and are usually more complex, or based on a different point-of-view.
The numerical results presented below include these extra algorithms, to justify empirically that despite their respective qualities, and despite being very efficient, it is reasonable to focus on \UCB{} and \klUCB{} in the rest of this thesis, as they stay comparable with or more efficient than other recent algorithms, but are simpler to implement (\eg, when comparing \UCB{} to \texttt{ApproximatedFHGittins} or $\mathrm{MOSS}$-$\mathrm{Anytime}$) to manipulate theoretically (\eg, when comparing \klUCB{} to BESA, PHE or RCB).

@@

$(ii)$
We detailed how SMPyBandits is implementing arms, problems, algorithms, and use these components to implement a simulation loop, with various visualizations being performed after the simulation.
As far as now, SMPyBandits is restricted to the finite-arm case, but it supports a wide range of arms distributions.
Different kinds of models are implemented, from stationary single-player to piece-wise stationary multi-players with different collision models.
One of the main qualities of the library is that it is quite exhaustive, as all the main families of algorithms covering these different models have been implemented, even very recent algorithms from the literature, as we tried our best in following the active research from December $2016$ to June $2019$.
More than $65$ algorithms or variants of algorithms are implemented for the single-player case, $5$ for the expert aggregation problem, about $15$ for the multi-players case, and about $20$ for the piece-wise stationary problem.
All the codebase is fully documented, and the library is using continuous integration to run automated tests on the code after every modification.
%
When comparing algorithms on a problem, the main performance measure is the regret, but the library also computes, stores and visualizes other measures, such as best-arm selection rate or mean cumulated reward, as well as real time and memory costs.
